Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic

### 2018-03-19 manuscript/manuscript.tex
added to conclusions section

The optimal statistic is particularly well-suited to compare multiple spatial correlation relations 
because using a different spatial correlation only requires changing the ORF 
in $\tilde{{\bf{S}}}_{IJ}$. 
\citet{thk+2016} demonstrated how the optimal statistic can be altered to fit for 
multiple spatial correlations at once in order to mitigate common noise sources such as 
clock error and ephemeris error. 
Here we compute the optimal statistic with monopole and dipole spatial correlations 
for the same simulated data sets as in the previous section in order to determine 
how well we can distinguish a GW background from a monopole or dipole signal. 
For a monopole signal, the ORF becomes simply
\[ \Gamma(\theta_{IJ}) = 1 \,\]
while for a dipole signal, the ORF becomes
\[ \Gamma(\theta_{IJ}) = \cos\theta_{IJ} \,. \]

@@

The noise-marginalized optimal statistic is a computationally inexpensive 
method for assessing the significance of an isotropic stochastic GWB in 
PTA data.

>>>>

The optimal statistic is particularly well-suited to compare multiple spatial correlation relations 
because using a different spatial correlation only requires changing the ORF 
in $\tilde{{\bf{S}}}_{IJ}$. 
\citet{thk+2016} demonstrated how the optimal statistic can be altered to fit for 
multiple spatial correlations at once in order to mitigate common noise sources such as 
clock error and ephemeris error. 
Here we compute the optimal statistic with monopole and dipole spatial correlations 
for the same simulated data sets as in the previous section in order to determine 
how well we can distinguish a GW background from a monopole or dipole signal. 
For a monopole signal, the ORF becomes simply
$\Gamma(\theta_{IJ}) = 1$, 
while for a dipole signal, the ORF becomes
$\Gamma(\theta_{IJ}) = \cos\theta_{IJ}$.

@@

The optimal statistic is a computationally inexpensive 
method for assessing the significance of an isotropic stochastic GWB in PTA data. 
The optimal statistic specifically looks for the spatial correlations characteristic 
of a GWB. These analyses are computationally expensive using a full Bayesian approach, 
but with the optimal statistic the significance of different spatial correlations 
can be computed very quickly. 
This makes the optimal statistic particularly well-suited for sky scramble analyses, 
where the ORF is scrambled many times to simulate randomizing the pulsars' sky locations.

However, the optimal statistic tends to systematically underestimate the strength of a GWB 
when computed for fixed values of the red noise because of 
covariances between red noise in individual pulsars and the GWB. 
In this paper, we introduce an improved method for computing the optimal statistic 
that marginalizes over the individual pulsars' red noises. 
This approach \sv{expand...}
